Growth Curves Paper

### 2020-03-18 R Analysis Scripts/CFU Data.Rmd
added comparisons of plate counts to the images I made with different thresholds

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(dplyr)
library(gdata)
library(ggplot2)
library(cowplot)
library(magrittr)
library(tidyr)
library(plotrix)
library(purrr)
library(moderndive)

#Data gathering functions:
GetGrowthCurveReplicate<- function (sample){
  bugCounts<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/GrowthCurves/", sample, ".csv"),header = TRUE)
  bugCounts<- na.omit(bugCounts)
  bugCounts<-bugCounts %>% mutate(logCFU = log10(CFU))
  bugCounts$sample<-sample
  return(bugCounts)
}
GetGrowthCurveCondition<-function(datalist){
  mappedData<-map(datalist, GetGrowthCurveReplicate)#this line reads all the files on the list
  boundData<-bind_rows(mappedData)
  colnames(boundData)[5] <- "strain"
  colnames(boundData)[8] <- "condition"
  return(boundData)
}
#Graphing functions
##this one graphs the log values
GraphGrowthCurveLog<-function (rawData, statsData){
    staticPlot<-ggplot()+
      geom_line(data=statsData%>% filter(bug == "Sa"& condition == "Static"), aes(x=time, y=logCFUmean, linetype = strain), size=1)+
      geom_errorbar(data=statsData%>% filter(bug == "Sa"& condition == "Static"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 1)+
      scale_linetype_manual(values = c( "dotted","solid"), labels = c("- HQNO", "wild type"))+
      labs(x = "Time (hours)", y = "log CFU", linetype = "Pa strain")+
      scale_y_continuous(limits = c(4,9))+
      theme_cowplot(12)

@@

```{r CFU data wrangling}
##remove stirred condition I didn't do in the scope and fix strain names:
DataComparisonCFU<-PA14DataStats %>% filter (condition == "Static") %>% select(bug, strain, CFUmean, logCFUmean, condition, time)%>% as.data.frame() 
DataComparisonCFU<-DataComparisonCFU%>% mutate(Strain = recode(DataComparisonCFU$strain, wt = "PA14wt", mut = "PA14mut"))

@@

##Adding more data from the TOB experiments!

#first select relevant conditions and fix column names
DataComparisonCFUSa<-newSCFMDataSa%>%  filter(!culture == "mono") %>% filter(antibiotic == "Control") %>% select(CFU, culture, bug, timePoint, antibiotic, logTranformed,sample, condition) %>% group_by(culture, timePoint) %>% summarise(meanCFU = mean(CFU), meanLogCFU = mean(logTranformed))

DataComparisonCFUSa<-DataComparisonCFUSa%>% as.data.frame() %>%mutate(Strain = recode(DataComparisonCFUSa$culture, wt = "PA14wt", mut = "PA14mut"))

colnames(DataComparisonCFUSa)<-c("strain", "time", "CFUs", "logCFUs", "Strain")

#then do the same for the scope data
DataComparisonScopeSa<-allConfocalResultsStats %>% filter(Species == "Sa")%>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,2,3,4)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0)) #The timePoint thing is because I start imaging after one hour and I don't have a time zero on scope
#merge them!
comparisonSaDf <- merge(DataComparisonCFUSa,DataComparisonScopeSa,by=c("time","Strain"))

@@

```{r graph CFU vs Confocal}
#Graph both Pa and Sa
##Original data as input:
ggplot()+
    geom_point(data=comparisonDf, aes(x=CFUs, y=MeanTotals,color = Species, shape = Strain),   size=4)+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))

@@

#Now we go to Sa

SaLinearModel <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Sa")%>% filter(!time == 0))
coefficients(SaLinearModel)
get_regression_summaries(SaLinearModel, print = TRUE)

ggplot()+
  geom_point(data=comparisonDf %>% filter(Species == "Sa"), aes(x=logCFUs,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =25.73, slope = -2.44, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 4.5, label = expression(paste("y=-2.44x+25.73  ", R^2, "= 0.619")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from CLSM")+
  theme_cowplot(12)

#Adding the early time points of Staph:
SaLinearModel2 <- lm(MeanLogTotals ~logCFUs, data = comparisonSaDf%>% filter(!time == 0))
coefficients(SaLinearModel2)
get_regression_summaries(SaLinearModel2, print = TRUE)
ggplot()+
  geom_point(data=comparisonSaDf %>% filter(Species == "Sa"), aes(x=logCFUs,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =-2.693, slope = 0.928, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 5.2, label = expression(paste("y=0.928x-2.693 ", R^2, "= 0.59")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from Scope")

>>>>

```{r setup, include=FALSE}
library(pixmap)
library(rtiff)
library(reshape2)
library(gplots)
library(dplyr)
library(gdata)
library(ggplot2)
library(cowplot)
library(magrittr)
library(tidyr)
library(plotrix)
library(purrr)
library(moderndive)
```
```{r data gathering functions}
#Data gathering functions:
GetGrowthCurveReplicate<- function (sample){
  bugCounts<-read.csv(paste0("/Users/jupabago/Documents/Whiteley/PROJECTS/PaSaProject/Spatial paper Data/GrowthCurves/", sample, ".csv"),header = TRUE)
  bugCounts<- na.omit(bugCounts)
  bugCounts<-bugCounts %>% mutate(logCFU = log10(CFU))
  bugCounts$sample<-sample
  return(bugCounts)
}
GetGrowthCurveCondition<-function(datalist){
  mappedData<-map(datalist, GetGrowthCurveReplicate)#this line reads all the files on the list
  boundData<-bind_rows(mappedData)
  colnames(boundData)[5] <- "strain"
  colnames(boundData)[8] <- "condition"
  return(boundData)
}
```

```{r graphing functions}
#Graphing functions

##this one graphs the log values
GraphGrowthCurveLog<-function (rawData, statsData){
    staticPlot<-ggplot()+
      geom_line(data=statsData%>% filter(bug == "Sa"& condition == "Static"), aes(x=time, y=logCFUmean, linetype = strain), size=1)+
      geom_errorbar(data=statsData%>% filter(bug == "Sa"& condition == "Static"), aes(x = time, ymin = logCFUmean-logCFUse, ymax = logCFUmean+logCFUse), width = 0, size = 1)+
      scale_linetype_manual(values = c( "dotted","solid"), labels = c("- HQNO", "wild type"))+
      labs(x = "Time (hours)", y = "log CFU", linetype = "Pa strain")+
      scale_y_continuous(limits = c(4,9))+
      theme_cowplot(12)

@@

```{r CFU data wrangling}

#GROWTH CURVES PLATE COUNTS DATA
##remove stirred condition I didn't do in the scope and fix strain names:
DataComparisonCFU<-PA14DataStats %>% filter (condition == "Static") %>% select(bug, strain, CFUmean, logCFUmean, condition, time)%>% as.data.frame() 
DataComparisonCFU<-DataComparisonCFU%>% mutate(Strain = recode(DataComparisonCFU$strain, wt = "PA14wt", mut = "PA14mut"))#add new column with proper name

@@

#FOR SA I can also use the Tob data, but first isolate from the other plate counts data:
DataComparisonCFUSaGrowthCurves<-DataComparisonCFU %>% filter(Species == "Sa") %>% select(Species, strain, CFUs, logCFUs, time, Strain)

#TOB PLATE COUNTS DATA
#this is dataframe is in the "SaPa TOB grwoth curves.RMD" file
#first select relevant conditions and fix column names
DataComparisonCFUSa<-newSCFMDataSa%>%  filter(!culture == "mono") %>% filter(antibiotic == "Control") %>% select(CFU, culture, bug, timePoint, antibiotic, logTranformed,sample, condition) %>% group_by(bug, culture, timePoint) %>% summarise(meanCFU = mean(CFU), meanLogCFU = mean(logTranformed)) %>% ungroup()

DataComparisonCFUSa<-DataComparisonCFUSa%>% mutate(Strain = recode(DataComparisonCFUSa$culture, wt = "PA14wt", mut = "PA14mut"))
colnames(DataComparisonCFUSa)<-c("Species", "strain", "time", "CFUs", "logCFUs", "Strain")

DataCFUAllSa<-bind_rows(DataComparisonCFUSa,DataComparisonCFUSaGrowthCurves)#combine growth curves and TOB
DataCFUAllSa<-DataCFUAllSa%>% group_by(Strain,time) %>% summarise(CFU = mean(CFUs), LogCFU = mean(logCFUs)) %>% ungroup()

#then do the same for the scope data
DataComparisonScopeSa<-allConfocalResultsStats %>% filter(Species == "Sa")%>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,2,3,4,7,11,15)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0)) #The timePoint thing is because I start imaging after one hour and I don't have a time zero on scope
#merge them!
comparisonSaDf <- merge(DataCFUAllSa,DataComparisonScopeSa,by=c("time","Strain"))
comparisonSaDfTOB <- merge(DataComparisonCFUSa,DataComparisonScopeSa,by=c("time","Strain"))

```
```{r microscopy data from different thresholds}
linear1Stats<-linear1[[2]] %>% filter(Species == "Sa")%>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,2,3,4,7,11,15)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0))

linear2Stats<-linear2[[2]] %>% filter(Species == "Sa")%>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,2,3,4,7,11,15)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0))

linear4Stats<-linear4[[2]] %>% filter(Species == "Sa")%>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,2,3,4,7,11,15)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0))

linear5Stats<-linear5[[2]] %>% filter(Species == "Sa")%>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,2,3,4,7,11,15)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0))

linear6Stats<-linear6[[2]] %>% filter(Species == "Sa")%>% filter (Strain %in% c("PA14wt","PA14mut")) %>% filter(Condition == "Pa:Sa 1:1") %>%  filter(TimePoint %in% c(0,2,3,4,7,11,15)) %>% select(Condition,TimePoint, Strain, Species, MeanTotals, MeanLogTotals) %>% mutate(time = ifelse(TimePoint>0, TimePoint+1, 0))

#merging with the whole growth curve data, not only TOB
comparisonSaDflinear1 <- merge(DataCFUAllSa,linear1Stats,by=c("time","Strain"))
comparisonSaDflinear2 <- merge(DataCFUAllSa,linear2Stats,by=c("time","Strain"))
comparisonSaDflinear4 <- merge(DataCFUAllSa,linear4Stats,by=c("time","Strain"))
comparisonSaDflinear5 <- merge(DataCFUAllSa,linear5Stats,by=c("time","Strain"))
comparisonSaDflinear6 <- merge(DataCFUAllSa,linear6Stats,by=c("time","Strain"))

@@

```{r graph CFU vs Confocal Pa}
#Graph both Pa and Sa
##Original data as input:
ggplot()+
    geom_point(data=comparisonDf, aes(x=CFUs, y=MeanTotals,color = Species, shape = Strain),   size=4)+
    scale_y_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))+
    scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x), labels = scales::trans_format("log10", scales::math_format(10^.x)))

@@

```{r graph CFU vs Confocal Sa}

#Now we go to Sa
#this is using only Growth cruves data
SaLinearModel <- lm(MeanLogTotals ~logCFUs, data = comparisonDf %>% filter (Species == "Sa")%>% filter(!time == 0))
coefficients(SaLinearModel)
get_regression_summaries(SaLinearModel, print = TRUE)

ggplot()+
  geom_point(data=comparisonDf %>% filter(Species == "Sa"), aes(x=logCFUs,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =25.73, slope = -2.44, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 4.5, label = expression(paste("y=-2.44x+25.73  ", R^2, "= 0.619")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from CLSM")+
  theme_cowplot(12)
#adding TOB data only
SaLinearModelTOB <- lm(MeanLogTotals ~logCFUs, data = comparisonSaDfTOB%>% filter(!time == 0))
coefficients(SaLinearModelTOB)
get_regression_summaries(SaLinearModelTOB, print = TRUE)
ggplot()+
  geom_point(data=comparisonSaDf, aes(x=LogCFU,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =20.243855, slope = -1.810558, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 5.2, label = expression(paste("y=-1.810+20.243855 ", R^2, "= 0.631")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from Scope")

#combining TOB and growth curves data
SaLinearModel2 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDf%>% filter(!time == 0))
coefficients(SaLinearModel2)
get_regression_summaries(SaLinearModel2, print = TRUE)
ggplot()+
  geom_point(data=comparisonSaDf, aes(x=LogCFU,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =20.243855, slope = -1.810558, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 5.2, label = expression(paste("y=-1.810+20.243855 ", R^2, "= 0.631")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from Scope")

SaLinearModel3 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDf%>% filter(!time %in% c(0,16)))
coefficients(SaLinearModel3)
get_regression_summaries(SaLinearModel3, print = TRUE)

SaLinearModel4 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDf%>% filter(!time %in% c(0,12,16)))
coefficients(SaLinearModel4)
get_regression_summaries(SaLinearModel4, print = TRUE)

SaLinearModel5 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDf%>% filter(!time %in% c(0,8,12,16)))
coefficients(SaLinearModel5)
get_regression_summaries(SaLinearModel5, print = TRUE)
ggplot()+
  geom_point(data=comparisonSaDf, aes(x=LogCFU,y=MeanLogTotals,  color = time, shape = Strain),size=4)+
  geom_abline(intercept =-.982, slope = .7210531, color = "red", size = 1, linetype = "dashed" )+
  scale_color_gradient(low="blue", high="red")+
  annotate("text", x = 7.9, y = 5.2, label = expression(paste("y=-1.810+20.243855 ", R^2, "= 0.631")), color = "red", size = 8)+
  labs(title =expression(paste(italic("S. aureus"), " CFU vs Confocal Biomass")), x = "Log CFUs",y = "Biomass from Scope")

#Testing the images with different thresholds

SaLinearModelLinear1 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDflinear1%>% filter(!time %in% c(0,12,16)))
coefficients(SaLinearModelLinear1)

SaLinearModelLinear2 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDflinear2%>% filter(!time %in% c(0,12,16)))
coefficients(SaLinearModelLinear2)

SaLinearModelLinear4 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDflinear4%>% filter(!time %in% c(0,12,16)))
coefficients(SaLinearModelLinear4)

SaLinearModelLinear5 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDflinear5%>% filter(!time %in% c(0,8,12,16)))
coefficients(SaLinearModelLinear5)

SaLinearModelLinear6 <- lm(MeanLogTotals ~LogCFU, data = comparisonSaDflinear6%>% filter(!time %in% c(0,8,12,16)))
coefficients(SaLinearModelLinear6)
